\documentclass[12pt]{article}

\usepackage[letterpaper, top=1in, bottom=1in, left=1in, right=1in]{geometry} % page
\usepackage[english]{babel} % language
\usepackage[utf8]{inputenc} % input
\usepackage[section]{placeins} % floats appear in the section they were defined in

\usepackage{
  amssymb, amsmath, % for math
  physics, % for physics notation
  siunitx, % for SI notation
  csquotes, % better biblatex
  listings, % for code
  graphicx, % for figures
  lipsum, % for blank text
  mathpazo, % for palatino font
  ragged2e, % for better alignment of text
  indentfirst, % indent the first line
  booktabs, % for tables
}

% PACKAGE SETTINGS
\allowdisplaybreaks % let the align environment span multiple pages
\setlength{\RaggedRightParindent}{\parindent} % fix ragged right

\usepackage[hidelinks]{hyperref}
\usepackage{cleveref}

% -----------------------------------------------------------------------------
% AUTHOR AND TITLE AND TITLEPAGE
% -----------------------------------------------------------------------------

\title{Environmental Data Analysis}
\author{Upmanu Lall}

\begin{document}
\maketitle
\RaggedRight

% -----------------------------------------------------------------------------
% START HERE
% -----------------------------------------------------------------------------


\section{Course Objectives}

\section{Course Outline}

\section{Course Organization}


\subsection{Course Communication}

Course files will be posted on the course \texttt{GitHub} page.
All class communication will be conducted through the Slack channel.
Please read the article at  \url{https://www.techrepublic.com/article/slack-goes-to-college-how-it-can-improve-the-classroom-experience/} to learn why we are using Slack for communication.
While the direct (private) messages are a great feature, please try to ask questions about course content in a public thread so that others can view and comment.

\subsection{Contact Information}

You can reach us at the information below:
\begin{tabular}{l l r}
  \toprule
  Name & Position & UNI \\
  \midrule
  Upmanu Lall & Professor & ula2 \\
  James Doss-Gollin & TA & jwd2136 \\
  \bottomrule
\end{tabular}

However, please use Slack for all course-related communication.


\section{Textbooks}

In this course we will not use any single textbook exhaustively, but will draw from the following textbooks.
Readings will be announced on the syllabus and in class.

\begin{itemize}
  \item Your first reference for anything related to programming, visualizing, or working with data in the R language is the R for Data Science book, available free online at \url{http://r4ds.had.co.nz/}.
  Reference this book for questions about reading data into R, using \texttt{Rmarkdown}, filtering and cleaning data, plotting data, dealing with dates, or dealing with strings.

  \item
\end{itemize}

% -----------------------------------------------------------------------------
% END HERE
% -----------------------------------------------------------------------------

\end{document}
